---
title: "HW6"
author: "Matt Untalan"
date: "2022-12-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Due date

Due: December 3 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |


### Problem 2

```{r clean}
homicide = 
  read_csv("data/homicide-data.csv", na = c("", "Unknown")) %>%
  mutate(
    city_state = paste(city, state, sep = ","),
    victim_age = as.numeric(victim_age),
    status = ifelse(disposition == "Closed by arrest", 1, 0), 
  ) %>%
  subset(!(city_state %in% c("Dallas,TX","Phoenix,AZ","Kansas City,MO","Tulsa,AL"))) %>%
  subset(victim_race == "White" | victim_race == "Black")
```

```{r balt}
glm_balt = 
  homicide %>%
  subset(city_state == "Baltimore,MD") %>%
  glm(status ~ victim_age + victim_sex + victim_race, data = ., family = binomial)

broom::tidy(glm_balt) %>% 
  mutate(
    OR = exp(estimate), 
    low_CI = exp(estimate - 1.96*std.error), 
    up_CI = exp(estimate + 1.96*std.error)
  ) %>%
  filter(term == "victim_sexMale") %>% 
  select(OR, low_CI, up_CI) %>% 
  knitr::kable(digits = 3)
```

The odds of a homicide being solved for a male victim is 0.426 times the odds of a homicide being solved for a female victim.  We are 95% confident that the true odds ratio falls between 0.325 and 0.558.

```{r all}
glm_all = 
  homicide %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    logit_solve = map(data, ~glm(status ~ victim_age + victim_sex + victim_race, family = binomial, data=.x)),
    logit_solve = map(logit_solve, broom::tidy)
  ) %>%
  select(-data) %>%
  unnest() %>%
  filter(term == "victim_sexMale") %>% 
  mutate(
    OR = exp(estimate), 
    low_CI = exp(estimate - 1.96*std.error), 
    up_CI = exp(estimate + 1.96*std.error), 
    city_state = fct_reorder(city_state, estimate)) %>% 
  select(city_state, OR, low_CI, up_CI)  
```

```{r graph}
glm_all %>%
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = low_CI, ymax = up_CI)) +
  coord_flip() +
  labs(title = "Odds Ratios of Homicides Solved for Male v. Female Victims",
       x = "City, State",
       y = "OR with CI")
```

Taking confidence interval into account, most cities solve appreciably less homicides with male victims than female victims.  In fact, no city shows an odds ratio of greater than 1 with a confidence interval that does not include 1.

### Problem 3

```{r recode}
birthweight = 
  read_csv("data/birthweight.csv")

birthweight_categ = 
  birthweight %>%
  mutate(
    babysex = recode(babysex, "1" = "male", "2" = "female"),
    babysex = factor(babysex, levels = c("male", "female")),
    frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
    frace = factor(frace, levels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    malform = recode(malform, "0" = "absent", "1" = "present"),
    malform = factor(malform, levels = c("absent", "present")),
    mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    mrace = factor(frace, levels = c("White", "Black", "Asian", "Puerto Rican", "Other"))
  )
```

Out of the variables provided, a study from the Journal of Epidemiology & Community Health mentions maternal age and smoking status as risk factors for low birthweight, with social factors such as family income.  Additionally, a 2017 Nature article analyzing birthweight across different races noted differences with significant covariates such as gestational weight gain and length of gestation.

```{r model}
birthweight_model = 
  lm(bwt ~ momage + smoken + fincome + mrace + wtgain + gaweeks, data = birthweight_categ)

birthweight_pred = 
  birthweight_categ %>%
  add_predictions(birthweight_model) %>%
  add_residuals(birthweight_model)

birthweight_pred %>%
  ggplot(aes(x = bwt, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(title = "Birthweight Resdiual Plot",
       x = "Birthweight (grams)",
       y = "Residual")
```

Creation of additional models for comparison and calculation of RMSE
```{r othermodels}
model_1 =
  lm(bwt ~ blength + gaweeks, data = birthweight_categ)

model_2 = 
  lm(bwt ~ bhead + blength + babysex + (bhead*blength) + (bhead*babysex) + (blength*babysex) + (bhead*blength*babysex), data = birthweight_categ)

cv_df = 
  crossv_mc(birthweight_categ, 100)

cv_df =
  cv_df %>%
  mutate(
    birthweight_model = map(train, ~lm(bwt ~ momage + smoken + fincome + mrace + wtgain + gaweeks, data = .x)),
    model_1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_2 = map(train, ~lm(bwt ~ bhead + blength + babysex + (bhead*blength) + (bhead*babysex) + (blength*babysex) + (bhead*blength*babysex), data = .x)) 
  ) %>%
  mutate(
    rmse_birthweight_model = map2_dbl(birthweight_model, test, ~rmse(model = .x, data = .y)),
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y))
  )

cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse, fill = model)) +
  geom_violin() +
  labs(title = "RMSE per Model",
       x = "Model",
       y = "RMSE")
```